---
output: html_document
---
lab rats

---
title: "The Relationship Between World Happiness Report Scores and WHO Country Suicide Rates - an Analysis"
author: "Lab rats"
output: github_document
---

```{r install-packages, echo=FALSE}
#install.packages("tidyverse")
library(tidyverse)

#library(janitor)

install.packages("naniar")
library (naniar)

install.packages("visdat")
library (visdat)

 install.packages ("leaflet")
library(leaflet) 
#
# install.packages("sf")
# library(sf) ## For spatial data
#
# install.packages("RColorBrewer")
# library(RColorBrewer) ## For colour palettes
#
# install.packages("htmltools")
# library(htmltools) ## For html
#
# install.packages("leafsync")
# library(leafsync) ## For placing plots side by side
#
# install.packages("kableExtra")
# library(kableExtra) ## Table output
#
# install.packages("stringr")
# library(stringr) ## String manipulation
#
# #install.packages ("rgeoboundaries")
# #library(rgeoboundaries)

#install.packages("gganimate")
#library(gganimate)
 
install.packages ("viridis")
library (viridis)
```

## Summary

Our group had great interest in public health. Thus, we engaged in the search for data focusing in public and mental health, finding sources from the World Health Organization (WHO) on global suicide rates from 1985 to 2021as the World Happiness Reports (WHR)  from 2015 to 2019. Since the WHR reports we found were bound to the years 2015-2019, we decided to extract the same years from the WHO world suicide dataset, to analyse if levels of happiness measured by the WHR parameters correlated with lower suicide rates. We were guided by the following question when we wwere formulating this project and our future visualizations: 

How does the score of a certain country on the World Happiness Report correlate with its respective suicide mortality rate?
As we continued our preliminary research, we found that taking a closer look at GDP per capita would also be enlightening for our findings. Initially, we also wanted to find the correlation between happiness, suicide rates, and the amount of light exposure received yearly in a certain country. However, after much research, we could not find any datasets that contained the average of light exposure by country and by year. Therefore, we added solely the question of relationship between GDP per capita and suicide rates as one of our research questions, framed as following: 
How does GDP per capita impact happiness, and to what extent does that data correlate with suicide mortality rates?


#### Cleaning data 
In order to make the visualizations we strived to do, we first had to make sure that both datasets we were working with followed the same standards for the categorization of countries and regions, and also that they represented information gathered in the same time frame. Initial data cleaning was composed of deletion of observations outside or our dataframe, as well as the deletion of columns that we deemed unnecessary for the purposes of this project. After discussion, our group decided that we should only use one metric for suicide rates – the suicide rates numeric – and that the country and region codes would not be necessary for the purpose of our visualizations. However, this latest assessment proved to be a mistake when tying to make a map visualization with package `leaflet`. To use `leaflet`, we had to merge our datasets with a shapefile dataset with international standards for country representation. Since the names of countries differed, the map did not show all of our data. This issue would have been easier to resolve if we had not deleted the columns of country codes for both the WHO suicide dataset and the WHR dataset, since country codes are unique to each country. 
Our process of cleaning and the failures and successes made us learn that sometimes, it is better to have more information than having less of it, because it allows for more room for later ideas. 


#### Visualizations and findings 
We wanted to make scaterplots to visualize the correlation of sucide rates and happiness scores and GDP per capita per country. As we continued making plots, we also wanted to visualize how the WHR parameters (perceptions of corruption, generosity, GDP per capita, social support, freedom and life expectancy)  influenced the overall happiness score of each country. We found that most of the WHR metrics show a positive correlation to the happiness scores, but GDP per capita seems to be the strongest. However, we were surprised to see that suicide rates and hapiness scores also had a positive correlation, showing that many countries that rank highly in the WHR report also have high numbers of sucide rates. The same was observed for GDP per capita and suicide rates – a positive correlation. 
Since the WHO suicide dataset was divided between sex categories and age groups, we also wanted to show some visualizaitons that considered sex categories and suicide rates. We made violin plots and bar plots to explore the gender difference of suicide rates. We found through the graphs that person identified as male do have much higher suicide rates across the globe if compared to persons that identify as female. We also found some outliers that either had too high suicide rates or too low, such as Lesotho and South Sudan. We conducted research to understand better the conditions on each country to try to understand what influenced their suicide rates beyond what was shown by the WHR report. 
Lastly, we wanted to make a cloropleth math showing the happiness scores by color gradient, and overlaying it with a bubble plot, with the size of the bubbles representing the suicide rates of each country. However, as stated in the previous question, we had stripped our datasets of crucial data to be able to succesfully merge it with a shapefile. Since we found that out too late in thee term we did not had enough time to comeback and retrieve some of the data lost to make a map visualization. We want to continue working on the map to see if we are capable to deliver this visualization. 


```{r load-data, include=FALSE}
whr_2015 <- read_csv("data/whr_2015.csv")
whr_2016 <- read_csv("data/whr_2016.csv")
whr_2017 <- read_csv("data/whr_2017.csv")
whr_2018 <- read_csv("data/whr_2018.csv")
whr_2019 <- read_csv("data/whr_2019.csv")
# suicide <- read_csv("data/suicide_2000_2019_v1.csv")


```

### Wrangling and Tidying data 

```{r tidy-data-2015, echo=FALSE, message=TRUE}
# whr_2015 <- whr_2015 %>%
#   mutate(Year = 2015) %>%
#   subset(select = -c(`Standard Error`,`Dystopia Residual`)) %>%
#    rename("Social Support" = "Family") %>%
#    rename("Perceptions of corruption" = "Trust (Government Corruption)") %>%
#   rename("Social support" = "Social Support") %>%
#     rename("GDP per capita" = "Economy (GDP per Capita)",
#           "Overall rank" = "Happiness Rank",
#           "Life expectancy" = "Health (Life Expectancy)") %>%
#    rename("Happiness score" = "Happiness Score") %>%
#  subset(select = -c(`Region`)) %>%
#   clean_names()
```

```{r reordering-cols-2015, echo=FALSE}

# whr_2015 <- whr_2015 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2016, echo=FALSE}
# whr_2016 <- whr_2016 %>%
# mutate(Year = 2016) %>%
#  subset(select = -c(`Lower Confidence Interval`, `Upper Confidence Interval`, `Dystopia Residual`)) %>%
#  rename("Social support"= "Family") %>%
#  rename("Perceptions of corruption" = "Trust (Government Corruption)") %>%
#  rename("Overall rank" = "Happiness Rank") %>%
#  rename("Happiness score" = "Happiness Score") %>%
#  rename("GDP per capita" = "Economy (GDP per Capita)") %>%
#  rename("Life expectancy" = "Health (Life Expectancy)") %>%
#  subset(select = -c(`Region`)) %>%
# clean_names()
```

```{r reordering-cols-2016 , echo=FALSE}
# whr_2016 <- whr_2016 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2017 , echo=FALSE}
# whr_2017 <- whr_2017 %>%
#    mutate(Year = 2017) %>%
#     subset(select = -c(`Whisker.high`,`Whisker.low`,`Dystopia.Residual`)) %>%
#        rename("Social Support"= "Family") %>%
#             rename("Overall rank" = "Happiness.Rank") %>%
#               rename("Happiness score" = "Happiness.Score") %>%
#               rename("GDP per capita" = "Economy..GDP.per.Capita.") %>%
#               rename("Life expectancy" = "Health..Life.Expectancy.") %>%
#               rename("Perceptions of corruption" = "Trust..Government.Corruption.") %>%
#   clean_names()

```

```{r reordering-cols-2017 , echo=FALSE}
# whr_2017 <- whr_2017 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2018 , echo=FALSE}
# whr_2018 <- whr_2018 %>% 
#   mutate (Year = 2018) %>%
#    rename("Freedom" = "Freedom to make life choices")%>%
#   rename("Life expectancy" = "Healthy life expectancy",
#          "Country" = "Country or region",
#          "Happiness score" = "Score") %>%
#   clean_names()
```

```{r reordering-cols-2018 , echo=FALSE}
# whr_2018 <- whr_2018 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2019, echo=FALSE}

whr_2019 <- whr_2019 %>%
   mutate (Year = 2019) %>%
  # rename("Freedom" = "Freedom to make life choices") %>%
  #  rename("Country" = "Country or region",
  #         "Life expectancy" = "Healthy life expectancy",
  #         "Happiness score" = "Score")%>%
   clean_names()
```

```{r reordering-cols-2019 , echo=FALSE}
# whr_2019 <- whr_2019 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r remove-col-2019, echo=FALSE}
 #whr_2019 <- subset(whr_2019, select = -c(year_2))
```

```{r expand-whr-dataset, echo=FALSE}
# whr_complete <- whr %>% 
#   expand(country,year)
# 
# whr_complete %>% dplyr::anti_join(whr)
# 
# whr <- whr %>% dplyr::right_join(whr_complete)

```


```{r create-whr , echo=FALSE}
# whr <- rbind(whr_2015,whr_2016, 
               # whr_2017, whr_2018, 
               # whr_2019) %>%
#   arrange(country)
```


```{r country-names-whr, echo=FALSE}
whr <- whr %>%
  mutate(country = case_when(country == "United States" ~ "USA",
                       country == "United Kingdom" ~ "UK",
                       country == "United Arab Emirates" ~ "UAE",
                       country == "Palestinian Territories" ~ "Palestine",
                       country == "Congo (Kinshasa)" ~ "DRC",
                       country == "Congo (Brazzaville)" ~ "Congo",
                       country == "Central African Republic" ~ "CAR",
     country == "Hong Kong S.A.R., China" ~ "Hong Kong", TRUE ~ country))

whr_2 <- subset(whr, !(country %in% c("North Cyprus", "Northern Cyprus", "Somaliland")))
```


```{r read-new-datasets, echo=FALSE}

whr <- read_csv("data/whr.csv")

whr_2 <- read_csv ("data/whr_2.csv")

write_csv( suicide, "suicide.csv")
# suicide <- read_csv ("data/suicide.csv")

suicide_2 <- read_csv("data/suicide_2.csv")
```


```{r suicide-tidy-data, echo=FALSE}
suicide <- suicide %>%
  clean_names() %>%
select(-c("who_region_code", "who_region", "country_code", "is_latest_year", "crude_suicide_rates_per_100_000_population_low_estimation", "crude_suicide_rates_per_100_000_population_high_estimation", "crude_suicide_rates_per_100_000_population_string")) %>%
rename("suicide_rates" = "crude_suicide_rates_per_100_000_population_numeric") %>%
filter(year %in% c("2015", "2016", "2017", "2018", "2019"))
```

```{r country-names-suicide, echo=FALSE}

suicide <- suicide %>%
mutate(country = case_when(country == "Syrian Arab Republic" ~ "Syria",
                           country == "United States of America" ~ "USA",
                           country == "United Republic of Tanzania" ~ "Tanzania",
                           country =="Russian Federation" ~ "Russia",
                           country == "The former Yugoslav Republic of Macedonia" ~ "North Macedonia",
                           country == "Lao People's Democratic Republic" ~ "Laos",
                           country == "Cabo Verde" ~ "Cape Verde",
                           country == "Brunei Darussalam" ~ "Brunei",
                           country == "Republic of Korea" ~ "South Korea",
                           country == "Democratic People's Republic of Korea" ~ "North Korea",
                           country == "Democratic Republic of the Congo" ~ "DRC",
                           country == "Iran (Islamic Republic of)" ~ "Iran",
                           country == "Republic of Moldova" ~ "Moldova",
                           country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
                           country == "United Arab Emirates" ~ "UAE",
                           country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
                           country == "Micronesia (Federated States of)" ~ "Micronesia",
                           country == "Bolivia (Plurinational State of)" ~ "Bolivia", TRUE ~ country))
```

```{r expand-suicide-dataset, echo=FALSE}

suicide_complete <- suicide %>% 
  expand (country, year, sex, age_group)

suicide_complete %>% dplyr::anti_join(suicide)

suicide_2 <- suicide %>% dplyr::right_join(suicide_complete)

```


## Visualizations

```{r whr-gdp-happiness, echo= FALSE, fig.alt= "This gif shows the correlation between GDP per capita and happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. That means higher GDP per capita correlates with higher happiness scores." }

gdp_per_capita_1 <- ggplot(whr_2, aes(x = gdp_per_capita, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "GDP per capita", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between GDP per capita\nand happiness score (WHR {frame_time})")

gdp_per_capita_2 <- animate(gdp_per_capita_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "gdp_per_capita_happiness.gif")
```


```{r whr-social-support-happiness, echo=FALSE , fig.alt= "This gif shows the correlation between access to social support and happiness score from 2015 through 2019 in the World Happiness Report. The two variables have a positive correlation. That means more social supports usually correlates with higher happiness scores."}

social_support_1 <- ggplot(whr_2, aes(x = social_support, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Social support", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between social support\nand happiness score (WHR {frame_time})")

social_support_2 <- animate(social_support_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "social_support_happiness.gif")
```

```{r whr-life-expectancy-happiness, echo=FALSE ,fig.alt= "This gif shows the correlation between life expectancy and happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. This means that longer life expectancies usually correlate with higher happiness scores."}

life_expectancy_1 <- ggplot(whr_2, aes(x = life_expectancy, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Life expectancy", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between life expectancy\nand happiness score (WHR {frame_time})")

life_expectancy_2 <- animate(life_expectancy_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "life_expectancy_happiness.gif")
```

```{r whr-freedom-happiness, echo= FALSE ,fig.alt= "This gif shows the correlation between freedom ratings and the happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. This means that more freedom usually correlates with higher happiness scores."}

freedom_1 <- ggplot(whr_2, aes(x = freedom, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Freedom", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between freedom\nand happiness score (WHR {frame_time})")

freedom_2 <- animate(freedom_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "freedom_happiness.gif")
```

```{r whr-generosity-happiness , echo= FALSE ,  fig.alt = "This gif shows the correlation between generosity percieved by each country and the happiness score in the World Happiness Report from 2015 through 2019. The two variables have a mild positive correlation. This means that more generosity does not frequently correlate with higher happiness scores."}

generosity_1 <- ggplot(whr_2, aes(x = generosity, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Generosity", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between generosity\nand happiness score (WHR {frame_time})")

generosity_2 <- animate(generosity_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "generosity_happiness.gif")
```

```{r whr-corruption-happiness, echo= FALSE , fig.alt= "This gif shows the correlation between the perceived level of corruption in each country and the happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. This means that less corruption correlates with higher happiness scores."}

perceptions_of_corruption_1 <- ggplot(whr_2, aes(x = as.numeric(perceptions_of_corruption), y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Perceptions of corruption", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    axis.title = element_text(size = 17),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between\nperceptions of corruption and\nhappiness score (WHR {frame_time})")

perceptions_of_corruption_2 <- animate(perceptions_of_corruption_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "perceptions_of_corruption_happiness.gif")
```

```{r missing-values, echo = FALSE , fig.alt= "This Bar Graph represents all possible missing data from our whr dataset. The X axis lists the variables that the data is organized under, and the Y axis represents number of observations. The red represents all character data, while the blue represents all numerical variables, The grey represents what variables are missing data inputs, which is 6.4% of all the data."}


#vis_miss(suicide_2)

# visdat::vis_dat(whr_2)
# 
# 
# visdat::vis_miss(whr_2)

```

```{r suicide-vs-happiness-scatter, echo = FALSE, fig.alt = "This faceted scatterplot looks at the correlation of three variables: suicide rates on the Y axis in regards to GDP on the X axis, and then WHR happiness scores that are being represented in a color gradient. Low happiness scores are expressed with deep purple and blue (3-4), mid-range scores are expressed with lighter blue and green (5-6), and high happiness scores are expressed with bright green and yellow (>=7)" fig.height=8, fig.width=15, warning=FALSE}

suicide_anim_1 <- suicide_2 %>%
  filter(sex == "Both sexes",
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(suicide_rates >= 0, happiness_score >= 0) %>%
  ggplot(aes(x = happiness_score, y = suicide_rates)) +
  geom_point(alpha = 0.5) +
  geom_miss_point(aes(color = happiness_score), alpha = 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "Happiness score", y = "Suicide\nrates", color = "WHR happiness\nscore",
       title = "Correlation between happiness score\nand suicide rates ({frame_time})") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 18),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.25),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 15)
  ) +
  scale_color_viridis_c() +
  ylim(0,40) +
  xlim(2.5,8) +
  transition_time(year)

suicide_gif <- animate(suicide_anim_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "suicide_happiness.gif")
```

```{r suicide-vs-GDP-scatter, echo = FALSE , fig.alt="This faceted scatterplot looks at the correlation of three variables: suicide rates on the Y axis in regards to GDP on the X axis, and then WHR happiness scores that are being represented in a color gradient. Low happiness scores are expressed with deep purple and blue (3-4), mid-range scores are expressed with lighter blue and green (5-6), and high happiness scores are expressed with bright green and yellow (>=7), fig.height=8, fig.width=15, warning=FALSE.}

suicide_anim <- suicide_2 %>%
  filter(sex == "Both sexes",
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(suicide_rates >= 0, happiness_score >= 0) %>%
  ggplot(aes(x = as.numeric(gdp_per_capita), y =  happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_miss_point(aes(color = happiness_score), alpha = 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "GDP per capita", y = "Suicide\nrates", color = "WHR happiness\nscore",
       title = "Correlation between GDP per capita,\nsuicide rates, and happiness score ({frame_time})") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 18),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.25),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 15)
  ) +
  scale_color_viridis_c() +
  ylim(0, 36) +
  transition_time(year)

suicide_gif <- animate(suicide_anim, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "suicide_gdp_happiness.gif")
```

```{r violin-plot, echo = FALSE, fig.alt="This violin plot shows the combined suicide rates of ramdomly selected countries according to sex and year. The countries were selected to represent each continent. The combined suicide rates of Males is seen in yellow; Females in purple. Female suicide rates are generally lower than Males, and Males have higher rates overall and a very high outlier. , fig.height=8, fig.width=15, warning=FALSE}

suicide_2 %>%
  filter(sex %in% c ("Male","Female"),
         age_group == "All Ages",
         country %in% c ("Lesotho","South Sudan",
                         "New Zealand","Portugal",
                         "China", "Brazil",
                         "Costa Rica", "Zimbabwe",
                         "Israel", "Mexico")) %>%
  mutate(year = factor(year, levels = c(2015, 2016, 2017, 2018, 2019)))%>%
  ggplot(aes(fill = sex, y = suicide_rates, x= year)) +
  geom_violin(position = "dodge", alpha = 0.5, outlier.colour = "transparent") +
  scale_fill_viridis_d(name = "Sex") +
  xlab("Year") +
  ylab("Suicide rates") +
  labs(title = "Collective suicide rates (per 100 000)\nof randomly selected countries",
       subtitle = "By sex") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 14),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 18, face = "bold", hjust = 0),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 13)
  )
```


```{r joined-datasets, echo=FALSE}

sui_joined_1 <- suicide_2 %>%
  mutate(country = case_when(country == "Syria" ~ "Syrian Arab Republic",
                           country == "USA" ~ "United States",
                           country == "Tanzania" ~ "United Republic of Tanzania",
                           country =="Russian Federation" ~ "Russia",
                           country == "North Macedonia" ~ "The former Yugoslav Republic of Macedonia",
                           country == "Laos" ~ "Lao People's Democratic Republic",
                           country == "Cape Verde" ~ "Cabo Verde",
                           country == "Brunei " ~ "Brunei Darussalam",
                           country == "South Korea" ~ " Republic of Korea",
                           country == "North Korea " ~ "Korea, Democratic People's Republic of",
                           country == "DRC" ~ "Democratic Republic of the Congo",
                           country == "Iran" ~ "Iran (Islamic Republic of)",
                           country == "Moldova" ~ " Republic of Moldova",
                           country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
                           country == "UAE" ~ "United Arab Emirates",
                           country == "UK" ~ "United Kingdom",
                           country == "Micronesia" ~ "Micronesia (Federated States of)",
                           country == "Bolivia" ~ "Bolivia (Plurinational State of)", TRUE ~ country)) %>%
   filter(sex == "Both sexes",
         age_group == "All Ages",
         year == 2019) %>%
  full_join(whr_2019, by = join_by(country, year)) %>%
  na.omit()
 
sui_joined_1$overall_rank <- as.numeric(as.character(sui_joined_1$overall_rank))
```


```{r map-viz-setup, echo=FALSE}

download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="data/world_shape_file.zip")

unzip("data/world_shape_file.zip")

 library(rgdal)
worldCountries <- readOGR(
  dsn= paste0(getwd(),"/data/TM_WORLD_BORDERS_SIMPL-0.3.shp") ,
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE)


merged_data <- merge(worldCountries, sui_joined_1, by.x = "NAME", by.y = "country", all.y = TRUE)


```

```{r list-color-palette , echo=FALSE}

viridis::inferno(256)

colors <- viridis::inferno(256)

```

```{r creating-palettes , echo=FALSE}

mypalette <- colorNumeric(palette= viridis(256, option = "B"), domain= merged_data$happiness_score, na.color="transparent")

mypalettesui <- colorNumeric(palette= viridis(256, option = "G"), domain= merged_data$suicide_rates, na.color="transparent")

```

```{r creating-map , echo=FALSE fig.alt = This graph is an atempt of showing the happiness scores of countries. It is a world map with a cloropleth visualization, with the nation boundaries colored with different colors according to their happiness score. In this visualization, shades of light yellow to orange are categorized as High Happiness Score, and shades of dark blue and purple represent low Happiness scores. Medium happiness scores are represented by red and pinkish tones. The graph is incomplete because inconsistencies in the dataset could not translate visually Therefore, some countries present no color. }
map <- merged_data %>% 
      leaflet() %>%
      addTiles() %>%
      setView( lat=10, lng=0 , zoom=1) %>% 
      setMaxBounds( lng1 = -180, lat1 = -87, 
                    lng2 = 200, lat2 = 97 ) %>%
  addPolygons(
      data = merged_data,
      fillColor = ~mypalette(happiness_score), 
      stroke=TRUE, 
      fillOpacity = 0.9,
      color="white", 
      weight=0.7, 
      # label = mytext, #label text variable
      # labelOptions = label, #label aesthetics variable
      # highlightOptions = highlight,
      group = "World Happiness Report") #group identifier
  #   )  %>%
  # addCircleMarkers(~long, ~lat, 
  #   fillColor = ~mypalettesui(suicide_rates), fillOpacity = 0.7, color="white", radius= suicide_rates, stroke=FALSE,
  #   label = mytext,
  #   labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto")
  # ) %>%
  # addLegend( pal=mypalette, values=~mag, opacity=0.9, title = "Magnitude", position = "bottomright" )

```

```{r map-legend1, echo=FALSE}
map <- map %>%
  addLegend(
    position = "bottomright",
    colors = c ("#F2F27DFF","#FBB61AFF","#F1731DFF","#CF4446FF","#85216BFF","#450A69FF","#000004FF"),
    values = ~merged_data$happiness_score,
    title = "World Happiness Score",
    label = c ("High score", "","", "Medium score", "","", "Low score"),
    opacity = 1)
 
map

```




```{r faceted-bar-plot , echo=FALSE, fig.alt = "This faceted bar plot is comparing all our randomly selected countries' suicide rate organized by sex. The countries are Zimbabwe (Africa), Mexico (N. America), Brazil (S. America), China (E. Asia), Israel (W. Asia), Costa Rica (C. America), Portugal (Europe), and New Zealand (Oceania). Sex is expressed with color, purple being female and yellow being male. Each country represents a similar trend of higher suicide rates of males compared to their female counterparts. We can also see that Zimbabwe has the highest rates of suicide overall compared to the other countries, but that there has been a steady decline since 2015."}

suicide_2 %>%
  filter(sex %in% c("Female", "Male") &
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(country %in% c("New Zealand", "Zimbabwe", "Portugal", "Brazil", "China", "Mexico", "Israel")) %>%
  group_by(year) %>%
  ggplot(aes(x = as.numeric(year), y= suicide_rates , fill = sex)) +
  geom_bar(position="dodge", stat="identity",) +
  scale_fill_viridis_d() +
  facet_wrap(~ country, ncol = 2) +
  coord_flip() +
  guides(y = guide_axis()) +
  labs(x = "Year",
       y = "Suicide rates",
       title = "Suicide rates (per 100 000)\nof randomly selected countries",
       subtitle = "Organized by sex",
       fill = "Sex") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 14),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 18, face = "bold", hjust = 0),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 13)
  )
```

```{r faceted-outliers, echo=FALSE fig.alt= These two bar plots are looking at the outliers of suicide rates on the highest (Lesotho) and lowest (South Sudan) points of the spectrum. The rates are organized by sex, yellow being make and purple being female. Suicide rate is the dependent variable on the Y axis and year is the variable on the X axis. South Sudan has a suicide rate of nearly zero for both sexes, while Lesotho has male suicide rates nearing 150 in 2015, with a steady decline in later years. Female suicide rates are substantially lower comparatively, with the highest rate in 2015 being around 40.}
suicide_2 %>%
   filter(sex %in% c("Female", "Male") &
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(country %in% c("Lesotho", "South Sudan")) %>%
  group_by(year) %>%
  ggplot(aes(x = as.numeric(year), y= suicide_rates , fill = sex)) +
  geom_bar(position="dodge", stat="identity",) +
  scale_fill_viridis_d() +
   facet_wrap(~ country, ncol = 2) +
  guides(y = guide_axis()) +
   labs(x = "Year",
       y = "Suicide Rates",
       title = "Suicide rates of outlier countries",
       fill = "Sex",
       subtitle = "Organized by sex") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 14),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 18, face = "bold", hjust = 0),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 13),
    strip.text.x = element_text(size = 13),
)
```


```{r south-sudan-barplot, echo=FALSE}
suicide_2 %>%
   filter(sex %in% c("Female", "Male") &
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(country == "South Sudan") %>%
  group_by(year) %>%
  ggplot(aes(x = year, y= suicide_rates , fill = sex)) +
  geom_bar(position="dodge", stat="identity", alpha = 0.7) +
  scale_fill_viridis_d() +
   labs(x = "Year",
       y = "Suicide Rates",
       title = "Suicide Rates of South Sudan By Year",
       subtitle = "Organized by Sex")
```
```



## Presentation

Our presentation can be found here:

https://docs.google.com/presentation/d/e/2PACX-1vTGLc4g4RlIvc3A5DvOiyx5sbPnOCC0YceVKR-iVQePDTLBCN3fi61DQfpZ4NY0-FS5JVHpIVyZ9spd/pub?start=false&loop=false&delayms=3000


```


## Data 


Mathurin Ache. (2021). World Happiness Report 2015-2021 [Data set]. Kaggle. https://www.kaggle.com/datasets/mathurinache/world-happiness-report-20152021?select=2015.csv Retrieved February 3, 2023.

Omkar Gowda. (2021). Suicide Rates Overview 1985 to 2021 [Data set]. Kaggle. https://www.kaggle.com/datasets/omkargowda/suicide-rates-overview-1985-to-2021 Retrieved February 3, 2023.


## References


Mathurin Ache. (2021). World Happiness Report 2015-2021 [Data set]. Kaggle. https://www.kaggle.com/datasets/mathurinache/world-happiness-report-20152021?select=2015.csv Retrieved February 3, 2023.

Omkar Gowda. (2021). Suicide Rates Overview 1985 to 2021 [Data set]. Kaggle. https://www.kaggle.com/datasets/omkargowda/suicide-rates-overview-1985-to-2021 Retrieved February 3, 2023.

World Bank. (2018). World development indicators: GDP (current US$) by country:1985 to 2016. Retrieved from http://databank.worldbank.org/data/source/world-development-indicators#

World Health Organization. (2018). Suicide prevention. Retrieved from http://www.who.int/mental_health/suicide-prevention/en/

Helliwell, John F., Richard Layard, Jeffrey Sachs, and Jan-Emmanuel De Neve, eds. 2020. World Happiness Report 2020. New York: Sustainable Development Solutions Network

Berghaus, R. (2008, October 6). Confronting Lesotho’s Health-Care Crisis. BU Today. https://www.bu.edu/articles/2009/confronting-lesotho%E2%80%99s-health-care-crisis/

