lab rats

---
title: "The Relationship Between World Happiness Report Scores and WHO Country Suicide Rates - an Analysis"
author: "Lab rats"
output: github_document
---

```{r install-packages }
#install.packages("tidyverse")
library(tidyverse)

#library(janitor)

install.packages("naniar")
library (naniar)

install.packages("visdat")
library (visdat)

 install.packages ("leaflet")
library(leaflet) 
#
# install.packages("sf")
# library(sf) ## For spatial data
#
# install.packages("RColorBrewer")
# library(RColorBrewer) ## For colour palettes
#
# install.packages("htmltools")
# library(htmltools) ## For html
#
# install.packages("leafsync")
# library(leafsync) ## For placing plots side by side
#
# install.packages("kableExtra")
# library(kableExtra) ## Table output
#
# install.packages("stringr")
# library(stringr) ## String manipulation
#
# #install.packages ("rgeoboundaries")
# #library(rgeoboundaries)

#install.packages("gganimate")
#library(gganimate)
 
install.packages ("viridis")
library (viridis)
```

## Summary

Write-up of your project and findings go here. 
Think of this as the text of your presentation. 
The length should be roughly 5 minutes when read out loud.
Although pacing varies, a 5-minute speech is roughly 750 words. 
To use the word count addin, select the text you want to count the words of (probably this is the 
Summary section of this document, go to Addins, and select the `Word count` addin).
This addin counts words using two different algorithms, but the results should be similar and as long as you're in the ballpark of 750 words, you're good! 
The addin will ignore code chunks and only count the words in prose.

You can also load your data here and present any analysis results / plots, but I strongly urge you to keep that to a minimum (maybe only the most important graphic, if you have one you can choose). 
And make sure to hide your code with  `echo = FALSE` unless the point you are trying to make is about the code itself.
Your results with proper output and graphics go in your presentation, this space is for a brief summary of your project.

```{r load-data, echo = FALSE}
whr_2015 <- read_csv("data/whr_2015.csv")
whr_2016 <- read_csv("data/whr_2016.csv")
whr_2017 <- read_csv("data/whr_2017.csv")
whr_2018 <- read_csv("data/whr_2018.csv")
whr_2019 <- read_csv("data/whr_2019.csv")
# suicide <- read_csv("data/suicide_2000_2019_v1.csv")


```

### Wrangling and Tidying data 

```{r tidy-data-2015}
# whr_2015 <- whr_2015 %>%
#   mutate(Year = 2015) %>%
#   subset(select = -c(`Standard Error`,`Dystopia Residual`)) %>%
#    rename("Social Support" = "Family") %>%
#    rename("Perceptions of corruption" = "Trust (Government Corruption)") %>%
#   rename("Social support" = "Social Support") %>%
#     rename("GDP per capita" = "Economy (GDP per Capita)",
#           "Overall rank" = "Happiness Rank",
#           "Life expectancy" = "Health (Life Expectancy)") %>%
#    rename("Happiness score" = "Happiness Score") %>%
#  subset(select = -c(`Region`)) %>%
#   clean_names()
```

```{r reordering-cols-2015}

# whr_2015 <- whr_2015 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2016}
# whr_2016 <- whr_2016 %>%
# mutate(Year = 2016) %>%
#  subset(select = -c(`Lower Confidence Interval`, `Upper Confidence Interval`, `Dystopia Residual`)) %>%
#  rename("Social support"= "Family") %>%
#  rename("Perceptions of corruption" = "Trust (Government Corruption)") %>%
#  rename("Overall rank" = "Happiness Rank") %>%
#  rename("Happiness score" = "Happiness Score") %>%
#  rename("GDP per capita" = "Economy (GDP per Capita)") %>%
#  rename("Life expectancy" = "Health (Life Expectancy)") %>%
#  subset(select = -c(`Region`)) %>%
# clean_names()
```

```{r reordering-cols-2016}
# whr_2016 <- whr_2016 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2017}
# whr_2017 <- whr_2017 %>%
#    mutate(Year = 2017) %>%
#     subset(select = -c(`Whisker.high`,`Whisker.low`,`Dystopia.Residual`)) %>%
#        rename("Social Support"= "Family") %>%
#             rename("Overall rank" = "Happiness.Rank") %>%
#               rename("Happiness score" = "Happiness.Score") %>%
#               rename("GDP per capita" = "Economy..GDP.per.Capita.") %>%
#               rename("Life expectancy" = "Health..Life.Expectancy.") %>%
#               rename("Perceptions of corruption" = "Trust..Government.Corruption.") %>%
#   clean_names()

```

```{r reordering-cols-2017}
# whr_2017 <- whr_2017 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2018}
# whr_2018 <- whr_2018 %>% 
#   mutate (Year = 2018) %>%
#    rename("Freedom" = "Freedom to make life choices")%>%
#   rename("Life expectancy" = "Healthy life expectancy",
#          "Country" = "Country or region",
#          "Happiness score" = "Score") %>%
#   clean_names()
```

```{r reordering-cols-2018}
# whr_2018 <- whr_2018 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r tidy-data-2019}

whr_2019 <- whr_2019 %>%
   mutate (Year = 2019) %>%
  # rename("Freedom" = "Freedom to make life choices") %>%
  #  rename("Country" = "Country or region",
  #         "Life expectancy" = "Healthy life expectancy",
  #         "Happiness score" = "Score")%>%
   clean_names()
```

```{r reordering-cols-2019}
# whr_2019 <- whr_2019 %>% 
#   select(c (overall_rank, country,happiness_score, gdp_per_capita,social_support, life_expectancy, freedom, generosity, perceptions_of_corruption, year))
```

```{r remove-col-2019}
 #whr_2019 <- subset(whr_2019, select = -c(year_2))
```

```{r expand-whr-dataset}
# whr_complete <- whr %>% 
#   expand(country,year)
# 
# whr_complete %>% dplyr::anti_join(whr)
# 
# whr <- whr %>% dplyr::right_join(whr_complete)

```


```{r create-whr}
# whr <- rbind(whr_2015, whr_2016, whr_2017, whr_2018, whr_2019) %>%
#   arrange(country)

# cut Somaliland, northern cyprus 
```


```{r country-names-whr}
whr <- whr %>%
  mutate(country = case_when(country == "United States" ~ "USA",
                       country == "United Kingdom" ~ "UK",
                       country == "United Arab Emirates" ~ "UAE",
                       country == "Palestinian Territories" ~ "Palestine",
                       country == "Congo (Kinshasa)" ~ "DRC",
                       country == "Congo (Brazzaville)" ~ "Congo",
                       country == "Central African Republic" ~ "CAR",
     country == "Hong Kong S.A.R., China" ~ "Hong Kong", TRUE ~ country))

whr_2 <- subset(whr, !(country %in% c("North Cyprus", "Northern Cyprus", "Somaliland")))
```


```{r read-new-datasets}

whr <- read_csv("data/whr.csv")

whr_2 <- read_csv ("data/whr_2.csv")

write_csv( suicide, "suicide.csv")
# suicide <- read_csv ("data/suicide.csv")

suicide_2 <- read_csv("data/suicide_2.csv")
```


```{r suicide-tidy-data}
suicide <- suicide %>%
  clean_names() %>%
select(-c("who_region_code", "who_region", "country_code", "is_latest_year", "crude_suicide_rates_per_100_000_population_low_estimation", "crude_suicide_rates_per_100_000_population_high_estimation", "crude_suicide_rates_per_100_000_population_string")) %>%
rename("suicide_rates" = "crude_suicide_rates_per_100_000_population_numeric") %>%
filter(year %in% c("2015", "2016", "2017", "2018", "2019"))
```

```{r country-names-suicide}

suicide <- suicide %>%
mutate(country = case_when(country == "Syrian Arab Republic" ~ "Syria",
                           country == "United States of America" ~ "USA",
                           country == "United Republic of Tanzania" ~ "Tanzania",
                           country =="Russian Federation" ~ "Russia",
                           country == "The former Yugoslav Republic of Macedonia" ~ "North Macedonia",
                           country == "Lao People's Democratic Republic" ~ "Laos",
                           country == "Cabo Verde" ~ "Cape Verde",
                           country == "Brunei Darussalam" ~ "Brunei",
                           country == "Republic of Korea" ~ "South Korea",
                           country == "Democratic People's Republic of Korea" ~ "North Korea",
                           country == "Democratic Republic of the Congo" ~ "DRC",
                           country == "Iran (Islamic Republic of)" ~ "Iran",
                           country == "Republic of Moldova" ~ "Moldova",
                           country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
                           country == "United Arab Emirates" ~ "UAE",
                           country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
                           country == "Micronesia (Federated States of)" ~ "Micronesia",
                           country == "Bolivia (Plurinational State of)" ~ "Bolivia", TRUE ~ country))
```

```{r expand-suicide-dataset}

suicide_complete <- suicide %>% 
  expand (country, year, sex, age_group)

suicide_complete %>% dplyr::anti_join(suicide)

suicide_2 <- suicide %>% dplyr::right_join(suicide_complete)

```


## Visualizations

```{r whr-gdp-happiness, fig.alt= "This gif shows the correlation between GDP per capita and happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. That means higher GDP per capita correlates with higher happiness scores." }

gdp_per_capita_1 <- ggplot(whr_2, aes(x = gdp_per_capita, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "GDP per capita", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between GDP per capita\nand happiness score (WHR {frame_time})")

gdp_per_capita_2 <- animate(gdp_per_capita_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "gdp_per_capita_happiness.gif")
```

```{r whr-social-support-happiness, fig.alt= ""This gif shows the correlation between access to social support and happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. That means more social supports usually correlates with higher happiness scores."}

social_support_1 <- ggplot(whr_2, aes(x = social_support, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Social support", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between social support\nand happiness score (WHR {frame_time})")

social_support_2 <- animate(social_support_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "social_support_happiness.gif")
```

```{r whr-life-expectancy-happiness, fig.alt= "This gif shows the correlation between life expectancy and happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. This means that longer life expectancies usually correlate with higher happiness scores."}

life_expectancy_1 <- ggplot(whr_2, aes(x = life_expectancy, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Life expectancy", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between life expectancy\nand happiness score (WHR {frame_time})")

life_expectancy_2 <- animate(life_expectancy_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "life_expectancy_happiness.gif")
```

```{r whr-freedom-happiness, fig.alt= "This gif shows the correlation between freedom ratings and the happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. This means that more freedom usually correlates with higher happiness scores."}

freedom_1 <- ggplot(whr_2, aes(x = freedom, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Freedom", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between freedom\nand happiness score (WHR {frame_time})")

freedom_2 <- animate(freedom_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "freedom_happiness.gif")
```

```{r whr-generosity-happiness fig.alt= "This gif shows the correlation between generosity percieved by each country and the happiness score in the World Happiness Report from 2015 through 2019. The two variables have a mild positive correlation. This means that more generosity does not frequently correlate with higher happiness scores."}

generosity_1 <- ggplot(whr_2, aes(x = generosity, y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Generosity", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 17),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between generosity\nand happiness score (WHR {frame_time})")

generosity_2 <- animate(generosity_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "generosity_happiness.gif")
```

```{r whr-corruption-happiness fig.alt= "This gif shows the correlation between the perceived level of corruption in each country and the happiness score in the World Happiness Report from 2015 through 2019. The two variables have a positive correlation. This means that less corruption correlates with higher happiness scores."}

perceptions_of_corruption_1 <- ggplot(whr_2, aes(x = as.numeric(perceptions_of_corruption), y = happiness_score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", aes(group = year)) +
  labs(x = "Perceptions of corruption", y = "Happiness\nscore") +
  labs(title = 'Year: {frame_time}') +
  transition_time(time = year) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    axis.title = element_text(size = 17),
    plot.background = element_rect(fill = "white")
  ) +
  ggtitle("The correlation between\nperceptions of corruption and\nhappiness score (WHR {frame_time})")

perceptions_of_corruption_2 <- animate(perceptions_of_corruption_1, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "perceptions_of_corruption_happiness.gif")
```

```{r missing-values fig.alt= "This Bar Graph represents all possible missing data from our whr dataset. The X axis lists the variables that the data is organized under, and the Y axis represents number of observations. The red represents all character data, while the blue represents all numerical variables, The grey represents what variables are missing data inputs, which is 6.4% of all the data."}


visdat::vis_miss(suicide_2)

visdat::vis_dat(whr)


visdat::vis_miss(whr)

```


```{r suicide-vs-GDP-scatter, fig.alt="This faceted scatterplot looks at the correlation of three variables: suicide rates on the Y axis in regards to GDP on the X axis, fig.height=8, fig.width=15, warning=FALSE, and then WHR happiness scores that are being represented in a color gradient. Low happiness scores are expressed with deep purple and blue (3-4), mid-range scores are expressed with lighter blue and green (5-6), and high happiness scores are expressed with bright green and yellow (>=7).}

suicide_anim <- suicide_2 %>%
  filter(sex == "Both sexes",
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(gdp_per_capita >= 0, suicide_rates >= 0, happiness_score >= 0) %>%
  ggplot(aes(x = as.numeric(gdp_per_capita), y = suicide_rates)) +
  geom_point(alpha = 0.5) +
  geom_miss_point(aes(color = happiness_score), alpha = 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "GDP per capita", y = "Suicide\nrates", color = "WHR happiness\nscore",
       title = "Correlation between GDP per capita,\nsuicide rates, and happiness score ({frame_time})") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 18),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.25),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 15)
  ) +
  scale_color_viridis_c() +
  ylim(0, 36) +
  transition_time(year)

suicide_gif <- animate(suicide_anim, nframes = 5, fps = 1.5, renderer = gifski_renderer())
anim_save(filename = "suicide_gdp_happiness.gif")
```

```{r violinplot, fig.alt="This violin plot shows the combined suicide rates of ramdomly selected countries according to sex and year. The countries were selected to represent each continent. The combined suicide rates of Males is seen in yellow; Females in purple. Female suicide rates are generally lower than Males, and Males have higher rates overall and a very high outlier. , fig.height=8, fig.width=15, warning=FALSE}

suicide_2 %>%
  filter(sex %in% c ("Male","Female"),
         age_group == "All Ages",
         country %in% c ("Lesotho","South Sudan",
                         "New Zealand","Portugal",
                         "China", "Brazil",
                         "Costa Rica", "Zimbabwe",
                         "Israel", "Mexico")) %>%
  mutate(year = factor(year, levels = c(2015, 2016, 2017, 2018, 2019)))%>%
  ggplot(aes(fill = sex, y = suicide_rates, x= year)) +
  geom_violin(position = "dodge", alpha = 0.5, outlier.colour = "transparent") +
  scale_fill_viridis_d(name = "Sex") +
  xlab("Year") +
  ylab("Suicide rates") +
  labs(title = "Collective suicide rates (per 100 000)\nof randomly selected countries",
       subtitle = "By sex") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 14),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 18, face = "bold", hjust = 0),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 13)
  )
```

```{r}
suicide_2 %>%
  filter(sex %in% c ("Male","Female"),
         age_group == "All Ages",
         country %in% c ("Lesotho","South Sudan",
                         "New Zealand","Portugal",
                         "China", "Brazil",
                         "Costa Rica", "Zimbabwe",
                         "Israel", "Mexico")) %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  mutate(country = factor(country, levels=c("Lesotho","SouthSudan",
                        "New Zealand","Portugal",
                         "China", "Brazil",
                         "Costa Rica", "Zimbabwe",
                         "Israel", "Mexico")))%>%
   ggplot(aes(fill=sex, y= suicide_rates, x= overall_rank)) +
  geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis_d(name= "Sex") +
    # theme_ipsum()  +
    xlab("Year") +
    ylab("Suicide rates") +
  labs(title = "Collective Suicide Rates of Randomly Selected Countries",
       subtitle = "By Sex") 
```

```{r finding-special-cases}
suicide_2 %>%
  filter(sex == "Both sexes", age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year))  %>%
  filter(gdp_per_capita <= 0.5, suicide_rates <= 4)
```

```{r joined-datasets}

sui_joined_1 <- suicide_2 %>%
  mutate(country = case_when(country == "Syria" ~ "Syrian Arab Republic",
                           country == "USA" ~ "United States",
                           country == "Tanzania" ~ "United Republic of Tanzania",
                           country =="Russian Federation" ~ "Russia",
                           country == "North Macedonia" ~ "The former Yugoslav Republic of Macedonia",
                           country == "Laos" ~ "Lao People's Democratic Republic",
                           country == "Cape Verde" ~ "Cabo Verde",
                           country == "Brunei " ~ "Brunei Darussalam",
                           country == "South Korea" ~ " Republic of Korea",
                           country == "North Korea " ~ "Korea, Democratic People's Republic of",
                           country == "DRC" ~ "Democratic Republic of the Congo",
                           country == "Iran" ~ "Iran (Islamic Republic of)",
                           country == "Moldova" ~ " Republic of Moldova",
                           country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
                           country == "UAE" ~ "United Arab Emirates",
                           country == "UK" ~ "United Kingdom",
                           country == "Micronesia" ~ "Micronesia (Federated States of)",
                           country == "Bolivia" ~ "Bolivia (Plurinational State of)", TRUE ~ country)) %>%
   filter(sex == "Both sexes",
         age_group == "All Ages",
         year == 2019) %>%
  full_join(whr_2019, by = join_by(country, year)) %>%
  na.omit()
 
sui_joined_1$overall_rank <- as.numeric(as.character(sui_joined_1$overall_rank))
```


```{r map-viz-setup}

# download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="data/world_shape_file.zip")
# 
# unzip("data/world_shape_file.zip")

 library(rgdal)
worldCountries <- readOGR(
  dsn= paste0(getwd(),"/TM_WORLD_BORDERS_SIMPL-0.3.shp") ,
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE)


merged_data <- merge(worldCountries, sui_joined_1, by.x = "NAME", by.y = "country", all.y = TRUE)

# worldCountries <- merge(sui_joined_1,by.x = "country", by.y = "NAME", all.x = TRUE)


```

```{r creating-map}
map <- merged_data %>% 
      leaflet() %>%
      addTiles() %>%
      setView( lat=10, lng=0 , zoom=1) %>% 
      setMaxBounds( lng1 = -180, lat1 = -87, 
                    lng2 = 200, lat2 = 97 ) %>%
  addPolygons(
      data = merged_data,
      fillColor = ~mypalette(happiness_score), 
      stroke=TRUE, 
      fillOpacity = 0.9,
      color="white", 
      weight=0.7, #size of lines
      # label = mytext, #label text variable
      # labelOptions = label, #label aesthetics variable
      # highlightOptions = highlight,
      group = "World Happiness Report") #group identifier
  #   )  %>%
  # addCircleMarkers(~long, ~lat, 
  #   fillColor = ~mypalettesui(suicide_rates), fillOpacity = 0.7, color="white", radius= suicide_rates, stroke=FALSE,
  #   label = mytext,
  #   labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto")
  # ) %>%
  # addLegend( pal=mypalette, values=~mag, opacity=0.9, title = "Magnitude", position = "bottomright" )

map
```

```{r map-legend1}
map <- map %>%
  addLegend(
    position = "bottomright",
    colors = c ("#F2F27DFF","#FBB61AFF","#F1731DFF","#CF4446FF","#85216BFF","#450A69FF","#000004FF"),
    values = ~merged_data$happiness_score,
    title = "World Happiness Score",
    label = c ("High score", "","", "Medium score", "","", "Low score"),
    opacity = 1)
 
map

```


```{r list-color-palette}

viridis::inferno(256)

colors <- viridis::inferno(256)

```

```{r creating-palettes}

mypalette <- colorNumeric(palette= viridis(256, option = "B"), domain= merged_data$happiness_score, na.color="transparent")

mypalettesui <- colorNumeric(palette= viridis(256, option = "G"), domain= merged_data$suicide_rates, na.color="transparent")

```


```{r faceted-bar-plot fig.alt = "This faceted bar plot is comparing all our randomly selected countries' suicide rate organized by sex. The countries are Zimbabwe (Africa), Mexico (N. America), Brazil (S. America), China (E. Asia), Israel (W. Asia), Costa Rica (C. America), Portugal (Europe), and New Zealand (Oceania). Sex is expressed with color, purple being female and yellow being male. Each country represents a similar trend of higher suicide rates of males compared to their female counterparts. We can also see that Zimbabwe has the highest rates of suicide overall compared to the other countries, but that there has been a steady decline since 2015.}

suicide_2 %>%
  filter(sex %in% c("Female", "Male") &
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(country %in% c("New Zealand", "Zimbabwe", "Portugal", "Brazil", "China", "Mexico", "Israel")) %>%
  group_by(year) %>%
  ggplot(aes(x = as.numeric(year), y= suicide_rates , fill = sex)) +
  geom_bar(position="dodge", stat="identity",) +
  scale_fill_viridis_d() +
  facet_wrap(~ country, ncol = 2) +
  coord_flip() +
  guides(y = guide_axis()) +
  labs(x = "Year",
       y = "Suicide rates",
       title = "Suicide rates (per 100 000)\nof randomly selected countries",
       subtitle = "Organized by sex",
       fill = "Sex") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 14),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 18, face = "bold", hjust = 0),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 13)
  )
```

```{r faceted-outliers}
suicide_2 %>%
   filter(sex %in% c("Female", "Male") &
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(country %in% c("Lesotho", "South Sudan")) %>%
  group_by(year) %>%
  ggplot(aes(x = as.numeric(year), y= suicide_rates , fill = sex)) +
  geom_bar(position="dodge", stat="identity",) +
  scale_fill_viridis_d() +
   facet_wrap(~ country, ncol = 2) +
  guides(y = guide_axis()) +
   labs(x = "Year",
       y = "Suicide Rates",
       title = "Suicide rates of outlier countries",
       fill = "Sex",
       subtitle = "Organized by sex") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 14),
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    plot.title = element_text(size = 18, face = "bold", hjust = 0),
    plot.background = element_rect(fill = "white"),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 13),
    strip.text.x = element_text(size = 13),
)
```


```{r south-sudan-barplot}
suicide_2 %>%
   filter(sex %in% c("Female", "Male") &
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(country == "South Sudan") %>%
  group_by(year) %>%
  ggplot(aes(x = year, y= suicide_rates , fill = sex)) +
  geom_bar(position="dodge", stat="identity", alpha = 0.7) +
  scale_fill_viridis_d() +
   labs(x = "Year",
       y = "Suicide Rates",
       title = "Suicide Rates of South Sudan By Year",
       subtitle = "Organized by Sex")
```
```

```{r lesotho-barplot}

suicide_2 %>%
   filter(sex %in% c("Female", "Male") &
         age_group == "All Ages") %>%
  full_join(whr_2, by = join_by(country, year)) %>%
  filter(country == "Lesotho") %>%
  group_by(year) %>%
  ggplot(aes(x = year, y= suicide_rates , fill = sex)) +
  geom_bar(position="dodge", stat="identity", alpha = 0.8) +
  scale_fill_viridis_d() +
   labs(x = "Year",
       y = "Suicide Rates",    
       title = "Suicide Rates of Lesotho By Year", 
       subtitle = "Organized by Sex")
       
```




## Presentation



```


## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
