---
title: "The Relationship Between World Happiness Report and WHO Country Suicide Rates - an Analysis"
author: "Lab rats"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)

```

```{r load-data}
#write_csv(whr, "whr.csv")
 whr <- read_csv("whr.csv")
# suicide <- read_csv("data/suicide_2000_2019_v1.csv")

```


## 1. Introduction
Our group had a common interest in public health and thought the topic would have ample opportunity for data analysis. 
We were able to find large data sets on suicide mortality rates from around the world, but wanted to know what factors contributed to higher or lower suicide rates in various countries.
We decided that comparing suicide rates to World Happiness Report (WHR) scores would allow us to see how the data narrates mental health for each country, and what specific variables show large contribution. 
WHR includes many independent variables, such as GDP per capita, family size, trust in government, generosity, etc.
We agreed that GDP per capita would allow us to see exactly how economy affects peoples' happiness, and therefore increase or decrease rates of happiness. 

Primary research question: How does the ranking of a certain country on the World Happiness Report correlate with their respective suicide mortality rate?
Secondary research question: How does GDP per capita impact happiness, and to what extent does that data correlate with suicide mortality rates?


## 2. Data

We chose two main datasets for our project. The first one was a compilation of suicide rates from the majority of countries created by the World Health Organization (WHO) 
for the period 2000 to 2019 (source: https://www.kaggle.com/datasets/mikekzan/who-crude-suicide-mortality-rate). 
The second one was a collection of scores of the WHR from 2015 to 2019, with data for each year located in a separate subdataset (source: https://www.kaggle.com/datasets/unsdsn/world-happiness).

However, we plan to modify these datasets to a significant degree to fit our needs. For example, we first intend to merge
the separate subdatasets of the WHR in one large dataset (with an additional column "year"). In addition, we want to harmonise the names of countries
all datasets (e.g., the WHO dataset has the "Syrian Arab Republic", whereas the WHR dataset has "Syria"). Afterwards, we will filter out countries that
do not appear in all datasets. Lastly, we will assign consistent regions for all countries, which will not be an easy endeavour,
recognising the inherent limitations of categorising countries with their unique characteristics.

Our first graph will be a scatterplot showing the correlation between the WHR scores and suicide mortality rates of all countries. Then we plan to
create such a graph for each region. Afterwards, we want to plot the correlation between GDP per capita (a variable in the WHR datasets) and the
happiness score on a global and regional scale. Lastly, we want to create graphs showing the correlation of the two aforementioned variables
with suicide mortality rates. However, it is of utmost importance to state that our plans are subject to change as the project evolves.

```{r glimpse}
# whr_2015 <- read_csv("data_copy/whr_2015.csv")
# whr_2016 <- read_csv("data_copy/whr_2016.csv")
# whr_2017 <- read_csv("data_copy/whr_2017.csv")
# whr_2018 <- read_csv("data_copy/whr_2018.csv")
# whr_2019 <- read_csv("data_copy/whr_2019.csv")
# whr <- rbind(whr_2015, whr_2016, whr_2017, whr_2018, whr_2019)
whr %>%
  glimpse() 
  
```


## 3. Data analysis plan
Variables we will consider for our analysis: 

```{r variables}

names(whr)

```


World Happiness Report dataset, for the years between 2015-2019: 
$ `Country`                         
$ `Region`                          
$ `Happiness Rank`                
$ `Happiness Score`                               
$ `Economy (GDP per Capita)`      
$ `Social Support`                         
$ `Health (Life Expectancy)`      
$ `Freedom`                         
$ `Trust (Government Corruption)` 
$ `Generosity `                     


World Suicide rates dataset, for the years between 2015-2019: 

Initially, we would also like to make an analysis associating light exposure to the happiness index and suicide rates. However, in our preliminary research so far we could not find a dataset with this information by country. We will continue the research to see if it is possible to add this variable from another dataset. 
=======

$`WHO Region Code`                                                
$ `WHO Region`                                                    
$ `Country Code`                                                  
$ `Country`                                                      
$ `Year`                                                         
$ `IsLatestYear`                                                 
$ `Sex`                                                           
$ `Age Group`                                                     
$ `Crude suicide rates (per 100 000 population) (numeric)`        
$ `Crude suicide rates (per 100 000 population) (low estimation)` 
$ `Crude suicide rates (per 100 000 population) (high estimation)`
$ `Crude suicide rates (per 100 000 population) (string)`

Initially, we would also like to make an analysis associanting light exposure to the happiness index and suicide rates. 
However, in our preliminary research so far we could not find a dataset with this information byt country. We will continue the research to see if it is possible to add this variable from another dataset. 

Since our dataset has a large set of observations, considering 155-158 countries out of the total countries in the world, we envision several possibilities to run our analysis. We could: 
* consider all of the countries in the dataset, grouping them by their regions 
* select a specific number of regions to analyze 
* equalize the population numbers for each region, trying to sample an equal amount of people per region/continent to be analyzed. Doing so, we would have to randomly exclude some countries of the analysis. 
*  choose specific countries to compare and contrast, dimishing considerably the number of observations. 
>>>>>>> 7a61a61afe8d6263d8900c921a4fafad95e325cf

